Servants, Syndicates, and Saviors: How Polarization and Prestige Impact the Political Behavior of Military Elites  (Prospectus)

### 2020-12-18 code/Erickson-dissprospect.v3.Rmd
commit

---
title: "Servants, Syndicates, and Saviors: The Rise of Domestic Political Interventionism by Senior Military Leaders (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation explores changes in how the contemporary military institution intervenes in domestic politics. I advance a theory which argues that normative beliefs of military officers regarding political activism is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility the military holds. To support my argument, I trace changes in levels and types of distinct forms of political activism by military officers, including authoring opinion-editorials in major newspapers, as domestic polarization and military institutional credibility rises. I ultimately argue that the increasing political activism of military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends involving the prestige of the military institution as well as political polarization that have been building for decades. I conclude by describing the implications of changing beliefs among military leaders regarding the role the military should have in domestic politics. This dissertation contributes to the fields of international relations and civil-military relations by advancing an empirical and historical analysis of a growing phenomenon: military leaders intervening in domestic American politics. _

@@

Across the world and in almost nearly every country, discussions regarding the participation of the military in politics are nothing new. Indeed, few topics evoke such strong opinions and emotions than whether and to what degree the military, both as a corporate body and as individual military officers, should be political vice remain separate from certain aspects of and activities associated with politics. These normative questions rightly form a significant portion of the study of civil-military relations, and intellectuals, practitioners, and pundits have frequently weighed in on these topics over the years.

Thankfully, within the United States, unlike many other countries across the world and throughout history, there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves. However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether. Indeed, a cursory examination of the relationships between Lincoln and McClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.

The central argument made in this dissertation is that increasing political activism at the hands of military officers indicates a shift in the beliefs of the military regarding its central role and purpose in society, or "role beliefs" [@fitch_armed_1998]. This shift, moreover, is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and the rising prestige and credibility of military institutions afforded by American society. The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient "servants" and towards that of "syndicates" or even "saviors" of the American state. This shift, moreover, holds enormous ramifications regarding the organizational culture of the military, its relationship with government and society, and American national security.

@@

My proposed dissertation topic features within the sub-field of civil-military relations, a broad academic sub-field that is generally placed within the International Relations and Comparative Politics sub-field (though within the American context, the sub-field of American Politics is entirely relevant and appropriate). Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and the relationship between a state's military and society.

At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or the question of how to ensure that a military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999]. Feaver describes this "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p. 214]. Moreover, the study of civil-military relations often focuses on the relationship between military and government elites, or at what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p. 380]. Indeed, it is at this level of interaction between government and military leaders that vital decisions are made, including decisions whether to go to war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.

A number of scholars have offered solutions to the"problematique" raised by Feaver, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, addresses this fundamental problem. Indeed, even more than 60 years after he wrote his book, Huntington's work often serves as the point of departure for many scholarly works on topics concerning civil-military relations. Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's normative conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes a concept that Huntington called "military professionalism" [@huntington_soldier_1957].

In _The Soldier and the State_, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States. The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3]. The societal imperative, Huntington claimed, was a further function of two distinct features of the United States. The first was an "ideological constant," which referred to America's political ideology of liberalism [@huntington_soldier_1957, 143-162]. Huntington argued that liberalism as a political ideology stressed "individualism" [@huntington_soldier_1957, 90], and that American society was, since its founding, suspicious of standing military organizations and professional military forces. The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143]. Huntington claimed that the tension generated by these two imperatives - one functional stemming from external threats, the other societal stemming from the character of American political ideology and structure - needed to be solved. Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society. Both goals - military security and civilian control - had to be obtained.

Figure 1, below, shows a visual diagram of Huntington's logic in a political science framework [@feaver_civil-military_1996]. As Peter Feaver points out in a 1996 article, Huntington did not use such language when developing and explaining his concept. According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military. Huntington viewed the key independent variable that enables the achievement of these goals while also resolving the tension generated by the opposing functional and societal imperatives as the adoption of a type of civil-military relations framework which he calls, "objective civilian control" of the military [@huntington_soldier_1957, 83]. In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and behavior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.

```{r figure one, include=FALSE, echo=FALSE}
huntington_dag <- dagify(
    MSCS ~ CC,
    CC ~ FI + SI,
    SI~I + S,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MSCS", 1, .5,
      "CC", .7, .5,
      "FI", .5, .6,
      "SI", .5, .4,
      "I", .3, .3,
      "S",.7, .3,
    ))

@@

Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83]. In fact, Huntington admits that the US Constitution actually envisioned and established a form of subjective vice objective scheme of civilian control of the military [@huntington_soldier_1957, 163-168]. What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats. It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general, that the military profession had emerged [@huntington_soldier_1957]. Huntington claims that Cold-War military officers, unlike the corporate body of officers that existed at the nation's founding, constitute a profession and thus, are part of a unique and distinct organization that differs from other occupations: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].

Numerous critiques surrounding the Huntingtonian conception of _military professionalism_ have been levied over the years. Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed. Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration - not segregation - of political and military activities. Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals. Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20]. Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020]. Risa Brooks charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs, which she alleges follows from a Huntingtonian conception of military professionalism, may develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17]. Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].

Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that an enduring issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020]. It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016]. In other words, the ground is always moving for the actors involved in civil-military relations.

# Framing the DV: Military Intervention in Politics and Military "Role Beliefs"

Given Huntington's normative model and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding the debates surrounding why the military would or should intervene in politics, and to what degree. Here again, the literature is vast. Scholars from the field of comparative politics have generally tackled the question of why militaries intervene in political matters, likely as a result of the variation that is available in country-specific or cross-national studies. It is vital to briefly understand a couple of these important frameworks.

The British political scientist, S.E. Finer, argues that military intervention in politics is ultimately a function of both the opportunity and desire to do so [@finer_man_1962]. Finer first identifies several factors that should inhibit the military from intervening in politics, including the concept of military professionalism (as espoused by Huntington), strict adherence to the principle of civilian supremacy, and the military's fear that intervention in politics would lead to it fracturing internally, and thus failing to remain a cohesive fighting force [@finer_man_1962]. Finer delivers a rather stinging critique of Huntingtonian military professionalism, and argues that although professionalism should work to focus the military on itself and thus, away from politics, the danger of the concept of military professionalism is that it can also be abused, particularly by military leaders who adopt the view that they serve the larger concept of "the state" rather than a particular civilian government [@finer_man_1962, 22-23]. Finer argues that military leaders throughout history, including von Seeckt and Gröner under the German Weimar Republic, and MacArthur under Truman in the United States, inappropriately intervened in politics precisely because their conceptions of "professionalism" demanded that they did so [@finer_man_1962], a critique of military professionalism that Risa Brooks shares [@brooks_paradoxes_2020]. Finer's whole point is that the principle of civilian control - not military professionalism - is and ought to serve as the only bulwark to keep the military out of politics.

@@

```{r figure two, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    RB ~ PC + PP + W,
    RB~PP,
    PC ~ PP,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "RB", 1, .5,
      "PC", .6, .5,
      "PP", .3, .7,
      "W", .8,.7,
    ))

@@

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("RB"="Military Role Beliefs", 
                     "PC"="Prestige and Credibility", 
                     "PP"="Political Polarization",
                     "W"="War")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="Determinants of US Military Role Beliefs")

@@

>>>>

---
title: "Servants, Syndicates, and Saviors: How Polarization and Prestige Impact the Political Behavior of Military Elites  (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation explores variation in the political behavior of the US military. The main theory that I advance is that across time, the political behavior of the US military is mainly impacted by two distinct factors: the level of domestic political polarization within a society, and the degree of prestige and credibility the military holds. These factors greatly influence the normative beliefs that militaries hold regarding their role in politics, and place at risk both the military's non-partisan ethic and the norm of civilian control of the military. As a result, the presence of these factors in domestic society impact the type, frequency, and other characteristics of political behaviors the military engages in. To support my theoretical argument, I conduct a mixed-methods empirical strategy that includes quantitative analysis and case study investigation. This dissertation advances scholarly understanding of how domestic politics impacts civil-military relations and international security._

@@

Across the world and in almost nearly every country, scholars, civilians, diplomats, and military leaders regularly discuss why and how militaries intervene in politics. In fact, it might be difficult to find a subject area other than civil-military relations where the practical behavior of soldiers and civilians, occurring before our very eyes, is so closely scrutinized and compared to normative conceptions of how we think these interactions between soldiers and statesmen should occur. For this reason, especially in the United States context, the political activity of the Armed Forces is a subject that has undergone intense scrutiny in recent years as polarization has, upon several occasions, thrust the military into the political spotlight.

From a now-retired Army Lieutenant Colonel providing Congressional testimony during President Donald Trump's impeachment proceedings, to a Navy Captain's removal for leaking a letter to the press that described his frustration with the military chain of command's response to the onset of the COVID-19 pandemic, to a remarkable panoply of retired military elites blasting the President's threat to use active duty forces to dispel protesters in cities across America during the summer of 2020, the events of the past few years have ignited an impassioned debate regarding the military's role in politics. Moreover, the near-ubiquitous use of militaries to respond to the COVID-19 pandemic has similarly awakened a desire to better understand the role of militaries in society, what they are used for, and how they should and actually do engage with civilian leaders. These healthy scholarly pursuits and debates, as well as the instances that justify their existence, will likely not end soon.

And while numerous scholarly works describe important facets of the political behavior of the military [@huntington_soldier_1957; @finer_man_1962; @fitch_armed_1998; @brooks_militaries_2009; @griffiths_not_2019; @brooks_paradoxes_2020], few if any fully describe how the presence of high degrees of political polarization and military prestige impact the manner in which the US military engages in distinct forms of political behavior, and importantly, how these forms of political behavior vary over time. This dissertation seeks to fill this important gap. In so doing, the intended audience groups of this dissertation are scholars, civilian officials, military officers, and (I hope) citizens who care about the connection between American society and its military.

The central argument made in this dissertation is that the conditions of high political polarization and a high degree of military prestige are present within a society, the result is an increasingly politically active military _if_ it chooses to be. This choice will largely depend on the motives and decisions of a small group of powerful military elites who, for better or for worse, are in charge of guiding the military through turbulent domestic waters. This dissertation shows that although the conditions of high polarization and military prestige have been prevalent within American society for the better part of the post-9/11 era, and especially in the past several years, this is not the first time these conditions have existed in the United States. Indeed, military elites have acted both nobly and dishonorably in the presence of these conditions at various time in the nation's past. It is important to learn from their mistakes so that they can be avoided again in the future, if possible.

This prospectus guides the reader in four parts. First, the prospectus introduces the reader to the general field of civil-military relations. In doing so, it emphasizes the importance of the dependent variable of study: political behavior of the military. Second, the prospectus briefly points out where the intended dissertation fits within the scholarly literature. Third, the prospectus lays out a theory of polarization and prestige and their impact on political behavior of the military. Fourth, the prospectus walks the reader through an empirical strategy for collecting the data required to test the theory laid out. In the conclusion, the prospectus addresses some immediate concerns and questions.

@@

The proposed dissertation topic regarding the political behavior by the military features within civil-military relations, a broad academic sub-field that is studied by scholars of International Relations, Comparative Politics, and American Politics (concerning the US military in particular). A broad range of scholarly interests comprises the study of civil-military relations, including coups, the role of domestic institutions, strategic effectiveness, the relationship between the military and society, and a host of other important questions.

At its root, the study of civil-military relations examines the question of how to ensure that a military is strong enough to defeat threats to the state without threatening the state itself. In the terminology of Peter Feaver, one prominent scholar, this question is indeed the "problematique" or central question within the field of all civil-military relations scholarship [@feaver_civil-military_1996; @feaver_civil-military_1999]. Feaver notes that this "problematique" is "vexing" and, actually a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p. 214].

The focus of study within civil-military relations can occur in several places, but as one might expect, there tend to be patterns depending on the specific area of scholarship. Sociologists studying the relationship between the military and society tend to focus on state-military relations, while political scientists often focus on the relationship between the state and militaries. Among the focus areas of political scientists is the relationships between prominent state officials and military leaders or elites. The civil-military relations scholar Risa Brooks, for instance, has rightly coined this relationship "leadership at the state's apex" [@brooks_risa_integrating_2019, p. 380], for it is at this level or set of interactions that government leaders and military officials make and implement vital decisions. These decisions include whether to go to war, how to implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.

A number of scholars have offered solutions, warnings, and caveats to the "problematique" Feaver describes - how the military can protect the state without threatening it - but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for a number of seminal debates within civil-military relations. Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's normative conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes "military professionalism" [@huntington_soldier_1957].

## Civilian Control of the Military and the Non-Partisan Military Ethic

In _The Soldier and the State_, Huntington asserts that liberal democracies such as the United States have to contend with, and ultimately balance, the simultaneous demands of military security and political liberalism. Huntington argued that in the aftermath of World War Two, American security requirements generated a "functional imperative" to maintain a military strong enough to defeat the  external threats of the Cold War, while the liberal character of American society generated a "societal imperative" such that the military as an institution could never become so alienated from society that it ceased to represent America [@huntington_soldier_1957, 1-3]. In making his argument, Huntington discusses a number of crucial differences between the military officer corps and the rest of American society. He claims that those who serve in the military hold to a distinct "mind" and "ethic" that views conflict as part of a "universal pattern throughout nature," and thus, that "the military view of man is...decidedly pessimistic" [@huntington_soldier_1957, 62-63]. On the other hand, American political liberalism, Huntington claims, is distinctly different. Huntington conceives of American liberal ideology as stressing "individualism," and that liberalism as an ideology "either denies the existence of power, minimizes its importance, or castigates it as inherently evil" [@huntington_soldier_1957, 90]. It is this juxtaposition between the unique character and ethic of militaries and the attributes of political liberalism in America, claims Huntington, that explains why America had been suspicious of standing military forces since its founding [@huntington_soldier_1957, 143-151].

In essence, Huntington is claiming that changes in the international security environment after World War Two are confronting the United States to adapt, whether it wants to or not, its relationship with its military. The United States had to learn, claims Huntington, to live with a sizable professional force capable of inflicting a significant degree of coercive power. At the same time, warns Huntington, America will not agree to such a relationship if its unique, liberal political ideology is threatened. Any solution to solving this tension must simultaneously render the military strong enough to defeat external threats, but completely subservient to civilian control.

Figure 1, below, displays the solution that Huntington puts forth to solve the tensions created in the strategic situation facing the US after World War Two.^[As Feaver points out in a 1996 article, Huntington did not use modern political science language when developing this concept. I constructed this puzzle using Feaver's 1996 recapitulation of Huntington.] The ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military. Huntington's proposed solution is to adopt a form of civilian control of the military known as "objective civilian control" in which civilian leaders grant the military a high degree of professional autonomy in exchange for the military staying out of partisan politics and instead, focusing on the professional expertise required to achieve American security [@huntington_soldier_1957, 83]. 
```{r figure one, include=FALSE, echo=FALSE}
huntington_dag <- dagify(
    MSCS ~ CC,
    CC ~ FI + SI,
    SI~I + S,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MSCS", 1, .5,
      "CC", .7, .5,
      "FI", .5, .6,
      "SI", .5, .4,
      "I", .3, .3,
      "S",.7, .3,
    ))

@@

```
The reason that such an arrangement would work, claimed Huntington, was because of the rise of the military as a profession that had unique corporate attributes. Unlike the military that had existed at the founding of America, which had been largely comprised of citizen-soldiers armed with small arms, innovation in wartime weapons and technologies had given rise to a professional body of officers skilled in what Harold Lasswell called the "management of violence" [Lasswell in @huntington_soldier_1957, 11].^[Of note, the topic of when the US military truly became a "profession" is a lively discussion. Historians tend to peg the dates at which the individual services embodied themselves as professions differently, ranging from...For more on that, see @need cite...] This was a development, Huntington claims, the American founders simply could not see. Because professions had unique corporate attributes, including that of developing a special body of knowledge and expertise, exercising a "social responsibility" by providing a vital service to society (defense), and fostering a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18], allowing the military a high degree of autonomy to exist qua profession would help ensure that it stay out of politics. It would be more interested in thriving as a profession than as a political body.

Numerous critiques surrounding the Huntingtonian conception of _military professionalism_ have been levied over the years. Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed. Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration, as opposed to segregation, of political and military activities. Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals. Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20].

Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020]. Risa Brooks, for instance, charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs may in practice develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17]. Her argument is simply that whether military leaders recognize it or not, military leaders comprise a profession that is irreducibly and inherently political to a certain extent. Rather than pretend this truth is not so, such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].

# Framing the DV: Variation in the Political Behavior of the Military

Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, there is nonetheless unity among scholars that in normative terms, civilian control of the military and a corresponding non-partisan ethic among military members are essential norms that should be observed and adhered to. Nonetheless, attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn constitutes a  [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020]. It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016]. In other words, the ground is always moving for the actors involved in civil-military relations.

Given Huntington's normative model and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding debates involving why and under what circumstances the military intervenes in politics. It is vital to briefly understand a couple of these important frameworks.

The British political scientist, S.E. Finer, argues that military intervention in politics is ultimately a function of the military having both the "opportunity" and "motive" to do so [@finer_man_1962]. Finer first identifies several factors that should inhibit the military from intervening in politics, including the concept of military professionalism (as espoused by Huntington), strict adherence to the principle of civilian supremacy, and the military's fear that intervention in politics would lead to it fracturing internally, and thus failing to remain a cohesive fighting force [@finer_man_1962]. Finer delivers a rather stinging critique of Huntingtonian military professionalism, and argues that although professionalism should work to focus the military on itself and thus, away from politics, the danger of the concept of military professionalism is that it can also be abused, particularly by military leaders who adopt the view that they serve the larger concept of "the state" rather than a particular civilian government [@finer_man_1962, 22-23]. Finer argues that military leaders throughout history, including von Seeckt and Gröner under the German Weimar Republic, and MacArthur under Truman in the United States, inappropriately intervened in politics precisely because their conceptions of "professionalism" demanded that they did so [@finer_man_1962], a critique of military professionalism that Risa Brooks shares [@brooks_paradoxes_2020]. Finer's whole point is that the principle of civilian control - not military professionalism - is and ought to serve as the only bulwark to keep the military out of politics.

@@

```{r figure two, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    PB ~ RB,
    RB ~ PC + PP + W,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "PB", 1,.5,
      "RB", .7, .5,
      "PC", .3, .7,
      "PP", .3, .5,
      "W", .3,.3,
    ))

@@

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("PB"="Political\n Behavior",
                     "RB"="Role Beliefs\n and Attitudes", 
                     "PC"="Prestige", 
                     "PP"="Polarization",
                     "W"="War")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="Determinants of Political Behavior by the Military")

@@

testimony of a now-retired Army Lieutenant Colonel  United States military  there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves. However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether. Indeed, a cursory examination of the relationships between Lincoln and McClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.

which they have done so.

understanding of how polarization and prestige impact the political behaviors of the military, and why.  as been the case for much of the post-9/11 era, but especially in the past several years, This increase in political activism is most concerning, moreover, among a small but powerful group of military elites who hold tremendously positive reputations among a large segment of the increasing political activism at the hands of military officers indicates a shift in the beliefs of the military regarding its central role and purpose in society, or "role beliefs" [@fitch_armed_1998]. This shift, moreover, is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and the rising prestige and credibility of military institutions afforded by American society. The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient "servants" and towards that of "syndicates" or even "saviors" of the American state. This shift, moreover, holds enormous ramifications regarding the organizational culture of the military, its relationship with government and society, and American national security.
